Hysteresis in human computation:\\ how one task affects another

### 2015-04-27 docs/publications/2015.04_PNAS/man.tex
EN edits to PNAS 2015.04

@@

We conducted variants of these experiments using different images, to see
whether this trend was robust.  In the experiment
\textit{intertask-food-culture}, workers were either assigned to a
\textit{food} or \textit{culture} treatment.  The initial tasks contained
either images depicting food (Fig.~\ref{fig:task}D), or depictions of cultural
scenes (of dance, sport, or music) (Fig.~\ref{fig:task}E).  The test tasks,
which were, again, identical for both treatments, depicted meals of
identifiable cultural origin (Fig.~\ref{fig:task}F).  As an aside, we recognize
that, despite efforts to the contrary, initial tasks in
\textit{intertask-food-culture} are arguably still of identifiable cultural
origin, but this fact would only tend to make our results more conservative (we
discuss this further in the supplementary text).

Results for this experiment again showed a strong bias as a result of intertask
effects (about 50\%) (Fig.~\ref{fig:theta}A).  Again, comparing intertask
effects to framing, the experiment (\textit{frame-food-culture}) used the same
test tasks as in \textit{intertask-food-culture}, but exposed workers to
statements framing the purpose of the tasks as the recognition of either food
or culture.  In this experiment, framing did not induce significant changes in
word frequencies (by $\chi^2$ test, $p=0.29$) (Fig.~\ref{fig:theta}A).

@@

This hypothesis is corroborated when we look at those words whose frequencies
changed the most from one treatment to another (Table~S\ref{table:top-words}).
The word ``food'', which is the most generic possible food-related word, was
always \textit{suppressed} among food-primed workers, relative to the
non-food-primed workers.  In fact, for all experiments, ``food'' was the
\textit{most suppressed} word.

@@

\section{Microtask setup} The work presented in the main text consisted of six
experiments involving HITs (bundles of tasks) performed on Amazon's Mechanical
Turk (MTurk) platform.  The flow of events experienced by a worker when
participating in this study is depicted in Fig.~S\ref{fig:task_schematic}.
Each experiment was divided into two treatments.  Treatments for a given
experiment differed with respect to a particular priming modality, which served
as the independent variable (as specified in Table~S\ref{table:experiments}).
The HITs for treatments of the same experiment always involved the same test
tasks (the last five tasks in the HIT), the results of which served as the
dependent variable.

All HITs were presented as a series of slides.  The first slide consisted of a
brief set of instructions.  For HITs that included framing, a framing statement
was shown on a slide immediately after the instructions (Fig.
S\ref{fig:hit_preamble}B and C).  Initial tasks, when included, followed next.
The initial tasks consisted of a series of five slides, each containing an
image-labeling task like the one shown in Fig.~S\ref{fig:hit_preamble}D.
Finally, for all HITs, five slides were shown containing test tasks, in the
same format as the initial tasks.

Workers were randomly assigned to the treatments, and a given worker could only
complete one HIT related to this study.  In \textit{intertask-food-culture},
the \textit{food} and \textit{culture} treatments were both divided into five
sub-treatments, in which the test tasks were permuted.  The permutations were
obtained by rotating the order shown in Fig.~S\ref{fig:task1:test} by different
amounts.  For example, rotating by one position involves moving the first image
to the second position, the second image to the third position, and so on, and
moving the last image to the first position.  The five different possible
rotations were used for the five sub-treatments.  This enabled us to study the
effects on each image when it occurred in different positions relative to the
initial tasks.

\section{Selection of images} The images used in the initial tasks for
\textit{intertask-food-objects}  are shown in Figs.~S\ref{fig:task1:food}
(\textit{food} treatment) and S\ref{fig:task1:obj} (\textit{objects}
treatment).  We took care to exclude objects from the food-images (except for
the dish supporting the food in some cases), and to exclude food from the
object-images.  The images used for the test tasks of
\textit{intertask-food-objects} and \textit{frame-food-objects} are shown in
Fig.~S\ref{fig:task1:test}.

The images used in the initial tasks for \textit{intertask-food-culture} are
shown in Figs.~S\ref{fig:task2:food} (\textit{food} treatment) and
S\ref{fig:task2:cult} (\textit{culture} treatment).  We selected the initial
images for the treatments of this experiment to respectively reflect depict
food and culture.  Naturally, since food is a very important aspect of culture,
pictorial depictions of the two concepts cannot be cleanly separated, and
images depicting food inevitably also depict culture.  However, the fact that
the images for the initial tasks in the \textit{food} treatment of
\textit{intertask-food-culture} also depict culture will only tend to make the
\textit{food} and \textit{culture} treatments more similar.  This would tend to
reduce the severity of intertask effects, but we nevertheless still observed
strong intertask effects.  Thus, the presence of culture in the food-related
initial images of \textit{intertask-food-culture} does not seem to have been
problematic.  The images used in the test tasks of
\textit{intertask-food-culture} and \textit{frame-food-culture} are shown in
Fig.~S\ref{fig:task2:test}.

\section{Framing} We framed tasks by including a message, near the beginning
of the HIT, that suggested a particular purpose for the tasks.  The precise
language used is shown in Table~S\ref{table:experiments}, and an example of a
framing slide is shown in Fig.~S\ref{fig:hit_preamble}B and C.  In
\textit{frame-food-objects} and \textit{frame-food-culture}, the framing
language was based on naming the (fictitious) entity running the study, for
example ``The Global Foundation for the Recognition of Cultures'', or
``Laboratory for the visual perception of Objects and Tools''.  In
\textit{echo-food-objects}, we stated the purpose explicitly (``The purpose of
this task is \dots''), and then required the worker to echo back the purpose of
the task by selecting it from a combo box Fig.~S\ref{fig:hit_preamble}C.

@@

>>>>

\usepackage{xr}
\externaldocument[S]{supp}
\usepackage{dcolumn}
\usepackage{multirow}
\usepackage{booktabs}
\newcolumntype{d}{D{.}{.}{4.0}}
\newcolumntype{s}{D{.}{.}{1.4}}

@@

We conducted variants of these experiments using different images, to see
whether this trend was robust.  In the experiment
\textit{intertask-food-culture}, workers were either assigned to a
\textit{food} or \textit{culture} treatment.  The initial tasks contained
either images depicting food (Fig.~\ref{fig:task}D), or depictions of cultural
scenes (of dance, sport, or music) (Fig.~\ref{fig:task}E).  The test tasks,
which were, again, identical for both treatments, depicted meals of
identifiable cultural origin (Fig.~\ref{fig:task}F).  As an aside, we recognize
that, despite efforts to the contrary, initial tasks in the food treatment of
\textit{intertask-food-culture} are arguably still of identifiable cultural
origin, but this fact would only tend to make our results more conservative (we
discuss this further in the supplementary text).

Results for this experiment again showed a strong bias as a result of intertask
effects (about 50\%) (Fig.~\ref{fig:theta}A).  Again, comparing intertask
effects to framing, the experiment (\textit{frame-food-culture}) used the same
test tasks as in \textit{intertask-food-culture}, but exposed workers to
statements framing the purpose of the tasks as the recognition of either food
or culture.  Again, we found that framing effects were weaker than 
intertask effects ($p=7.6\times10^{-7}$).  In fact, framing did not induce 
significant changes in word frequencies (by $\chi^2$ test, $p=0.29$) 
(Fig.~\ref{fig:theta}A).

@@

This hypothesis is corroborated when we look at those words whose frequencies
changed the most from one treatment to another (Table~\ref{table:top-words}).
As one would expect, references to objects, such as ``candle'', ``fork'', 
or ``statue'' appear less frequently among the labels provided by workers in 
food-exposed treatments.  But, the food-exposed workers also provide the most
generic references to to food less often, such as the word ``food'' itself, 
or ``meal''.  
In fact, the word ``food'', which is the most generic possible food-related 
word, was always \textit{the most suppressed} among food-primed workers, 
relative to their non-food-primed counterparts.  
Meanwhile, food-exposed workers provide more specific references to food more 
often, such as ``coffee'', ``cheese'', or ``wine''.

\begin{table*}
	\centering
	\setlength{\tabcolsep}{4pt}
	\begin{tabular}{ c c c c c }

		\setlength{\tabcolsep}{4pt}
		\begin{tabular}{ r | c }
		\toprule
		\multicolumn{2}{c}{
			\parbox[c]{2.5cm}{
				\centering
					\textit{intertask-food-objects}
			}} \\
		\midrule
		coffee & 38 \\
		meal & 34 \\
		cheese & 34 \\
		apple & 32 \\
		dessert & 21 \\
		cup & -30 \\
		glass & -45 \\
		table & -70 \\
		candle & -74 \\
		food & -80 \\
		\bottomrule
		\end{tabular}

&

		\setlength{\tabcolsep}{4pt}
		\begin{tabular}{ r | c }
		\toprule
		\multicolumn{2}{c}{
			\parbox[c]{2.5cm}{
				\centering
				\textit{frame-food-objects}
			}
		}\\
		\midrule
		bread & 18 \\
		wine & 18 \\
		cheese & 16 \\
		apple & 14 \\
		oil & 12 \\
		table & -9 \\
		meal & -10 \\
		candle & -12 \\
		dinner & -13 \\
		food & -32 \\
		\bottomrule
		\end{tabular}

&

		\setlength{\tabcolsep}{4pt}
		\begin{tabular}{ r | c }
		\toprule
		\multicolumn{2}{c}{
			\parbox[c]{2.3cm}{
				\centering
			\textit{echo-food-objects}} 
		}\\
		\midrule
		apple & 24 \\
		cheese & 23 \\
		wine & 15 \\
		coffee & 14 \\
		oil & 7 \\
		knife & -24 \\
		dinner & -26 \\
		fork & -27 \\
		candle & -35 \\
		food & -55 \\
		\bottomrule
		\end{tabular}

&

		\setlength{\tabcolsep}{4pt}
		\begin{tabular}{ r | c }
		\toprule
		\multicolumn{2}{c}{
			\parbox[c]{2.5cm}{
				\centering
			\textit{intertask-food-culture}} 
			\smallskip
		}\\

		\midrule
		spicy & 26 \\
		sauce & 17 \\
		indian & 15 \\
		buffet & 14 \\
		exotic & 12 \\
		festival & -11 \\
		offering & -12 \\
		statue & -15 \\
		india & -20 \\
		food & -56 \\
		\bottomrule
		\end{tabular}

&

		\setlength{\tabcolsep}{4pt}
		\begin{tabular}{ r | c }
		\toprule
		\multicolumn{2}{c}{
			\parbox[c]{2.5cm}{
				\centering
			\textit{frame-food-culture}} 
		}\\
		\noalign{\smallskip}
		\midrule
		indian & 11 \\
		banquet & 8 \\
		spicy & 7 \\
		asian & 6 \\
		variety & 6 \\
		delicious & -6 \\
		meat & -7 \\
		festival & -7 \\
		spice & -7 \\
		food & -9 \\
		\bottomrule
		\end{tabular}

	\end{tabular}
	\caption{
		The five words whose frequencies increased, and those whose 
		frequencies decreased, the most, between treatments of given 
		experiments, within labels attributed to the first test task.
		Positive values indicate that the food-exposed treatment
		used the word with higher frequency.  
		Note that the word ``food'' is always the most suppressed among
		food-exposed workers.
		Word frequencies for 
		\textit{intertask-food-objects} correspond to the labels attributed
		to image 1 (see Fig. \ref{Sfig:task1:test}) which is not necessarily 
		the first 
		test task due to the tasks being permuted.  There were five times
		as many workers in \textit{intertask-food-objects}, hence
		larger absolute differences are observed.
	}
	\label{table:top-words}
\end{table*}

@@

\section{Microtask setup} The work presented in the main text consisted of six
experiments involving HITs (bundles of tasks) performed on Amazon's Mechanical
Turk (MTurk) platform.  The flow of events experienced by a worker when
participating in this study is depicted in Fig.~S\ref{Sfig:task_schematic}.
Each experiment was divided into two treatments.  Treatments for a given
experiment differed with respect to a particular priming modality, which served
as the independent variable (as specified in Table~S\ref{Stable:experiments}).
The HITs for treatments of the same experiment always involved the same test
tasks (the last five tasks in the HIT), the results of which served as the
dependent variable.

All HITs were presented as a series of slides.  The first slide consisted of a
brief set of instructions.  For HITs that included framing, a framing statement
was shown on a slide immediately after the instructions (Fig.
S\ref{Sfig:hit_preamble}B and C).  Initial tasks, when included, followed next.
The initial tasks consisted of a series of five slides, each containing an
image-labeling task like the one shown in Fig.~S\ref{Sfig:hit_preamble}D.
Finally, for all HITs, five slides were shown containing test tasks, in the
same format as the initial tasks.

Workers were randomly assigned to the treatments, and a given worker could only
complete one HIT related to this study.  In \textit{intertask-food-culture},
the \textit{food} and \textit{culture} treatments were both divided into five
sub-treatments, in which the test tasks were permuted.  The permutations were
obtained by rotating the order shown in Fig.~S\ref{Sfig:task1:test} by different
amounts.  For example, rotating by one position involves moving the first image
to the second position, the second image to the third position, and so on, and
moving the last image to the first position.  The five different possible
rotations were used for the five sub-treatments.  This enabled us to study the
effects on each image when it occurred in different positions relative to the
initial tasks.

\section{Selection of images} The images used in the initial tasks for
\textit{intertask-food-objects}  are shown in Figs.~S\ref{Sfig:task1:food}
(\textit{food} treatment) and S\ref{Sfig:task1:obj} (\textit{objects}
treatment).  We took care to exclude objects from the food-images (except for
the dish supporting the food in some cases), and to exclude food from the
object-images.  The images used for the test tasks of
\textit{intertask-food-objects} and \textit{frame-food-objects} are shown in
Fig.~S\ref{Sfig:task1:test}.

The images used in the initial tasks for \textit{intertask-food-culture} are
shown in Figs.~S\ref{Sfig:task2:food} (\textit{food} treatment) and
S\ref{Sfig:task2:cult} (\textit{culture} treatment).  We selected the initial
images for the treatments of this experiment to respectively depict
food and culture.  Naturally, since food is a very important aspect of culture,
pictorial depictions of the two concepts cannot be cleanly separated, and
images depicting food inevitably also depict culture.  However, the fact that
the images for the initial tasks in the \textit{food} treatment of
\textit{intertask-food-culture} also depict culture will only tend to make the
\textit{food} and \textit{culture} treatments more similar.  This would tend to
reduce the severity of intertask effects, but we nevertheless still observed
strong intertask effects.  Thus, the presence of culture in the food-related
initial images of \textit{intertask-food-culture} does not seem to have been
problematic.  The images used in the test tasks of
\textit{intertask-food-culture} and \textit{frame-food-culture} are shown in
Fig.~S\ref{Sfig:task2:test}.

\section{Framing} We framed tasks by including a message, near the beginning
of the HIT, that suggested a particular purpose for the tasks.  The precise
language used is shown in Table~S\ref{Stable:experiments}, and an example of a
framing slide is shown in Fig.~S\ref{Sfig:hit_preamble}B and C.  In
\textit{frame-food-objects} and \textit{frame-food-culture}, the framing
language was based on naming the (fictitious) entity running the study, for
example ``The Global Foundation for the Recognition of Cultures'', or
``Laboratory for the visual perception of Objects and Tools''.  In
\textit{echo-food-objects}, we stated the purpose explicitly (``The purpose of
this task is \dots''), and then required the worker to echo back the purpose of
the task by selecting it from a combo box Fig.~S\ref{Sfig:hit_preamble}C.

@@

\section{Derivation of bias inequality}
A sketch of the derivation of the lower bound on the degree of bias 
$\theta$ in terms of the observed accuracy of a classifier, $\eta$, follows
(see the supplementary material for a detailed derivation). 
Consider the accuracy of the best possible classifier, $\eta^*$.  
Assume two groups of workers, have been
exposed to different treatments, 1 and 2, and as a result have
a different probability of yielding response $x$ during a task: 
$p_1(x)$ and $p_2(x)$.  Upon observing response $x$, it is optimal to guess 
that the worker came from the treatment that produces response $x$ more 
often: i.e. to guess $z=\arg\max_z(p_z(x))$.

The probability that this guess is correct is:
\begin{align}
	\Pr\{\mathrm{correct}\vert x\} 	&= \frac{1}{2} + \frac{|p_0(x') - p_1(x')|}{2 \big(p_0(x') + p_1(x') \big)}.
\end{align}
The probability that the optimal classifier will be correct in general,
which is by definition $\eta^*$, can be found by summing over all values of 
$x$:

\begin{align}
	\eta^* &= \sum_{x' \in \mathcal{X}} \mathrm{Pr}\{\mathrm{correct}\vert x'\}
			\mathrm{Pr}\{x'\} \\
		&= \sum_{x'\in X} \left(
				\frac{1}{2} + 
				\frac{|p_0(x') - p_1(x')|}{2 \big(p_0(x') + p_1(x') \big)}
			\right)
			\left(
				\frac{p_0(x') + p_1(x')}{2}
			\right)\\
		&= \frac{1 + \theta}{2}\\
		\implies \theta &= 2\eta^* - 1.
\end{align}

Since no classifier can be more accurate than the optimal classifier, we 
obtain the bound stated in Eq.~\ref{l1}.
